HMMotion: Using tracking data to predict coverage strategies in American Football

### 2025-01-06 submission.Rmd

We compare the predictive performance of our three models to determine the effectiveness of player movements before snap for detecting the correct defensive scheme. We train the models on 85% of the data and evaluate performance on the remaining 15%. Since we use cross-validation for tuning of the models during the training phase, we perform our evaluation repeatedly. We use 50 repetition each time splitting the data differently, and for each we tune and train a model and then evaluate the performance on the test set. We do this to ensure that our results are meaningful and not due to chance (especially in our small data set).

@@

<div style="text-align: center;">
  ![](ACCs_bdb_main.png)
  ![](AUCs_bdb_main.png)
</div>

@@

>>>>

<!--
We compare the predictive performance of our three models to determine the effectiveness of player movements before snap for detecting the correct defensive scheme. We train the models on 85% of the data and evaluate performance on the remaining 15%. Since we use cross-validation for tuning of the models during the training phase, we perform our evaluation repeatedly. We use 50 repetition each time splitting the data differently, and for each we tune and train a model and then evaluate the performance on the test set. We do this to ensure that our results are meaningful and not due to chance (especially in our small data set).
-->

`r colorize("Vorschlag (Robert):","red")`

We evaluate the predictive performance of our three models to assess how well pre-snap player movements help identify the correct defensive scheme. To do this, we split the data into two sets: 85% for training and 15% for testing. During the training phase, we use cross-validation to fine-tune the models.

To ensure robust and reliable results, we repeat the evaluation process 50 times, using different cv splits for each iteration. For each split, we train and fine-tune the model on the training data and then evaluate its performance on the test set. This approach helps minimize the risk of our findings being influenced by randomness, which is particularly important given the limited size of our dataset.

@@

`r colorize("Robert: komisch, bei mir sind sie eig recht scharf... hab versucht es bisschen zu ändern. Wenns das nicht besser macht, weiß ich leider auch nich was hilft...","red")`

<div style="text-align: center;">
  ![](ACCs_bdb_main2.png)
  ![](AUCs_bdb_main2.png)
</div>

@@

`r colorize("Vorschlag Robert (mit Teilen von oben und chatGPT Hilfe von unten):","red")`

This project explored the use of hidden Markov models (HMMs) to enhance the prediction of defensive schemes in football by incorporating pre-snap player movements. Models augmented with HMM-derived features demonstrated improved predictive performance compared to those relying solely on naive motion data.

Comparing predictions from pre-motion and post-motion models provides insights into an offense's ability to use motion effectively to identify defensive coverage schemes. Similarly, it allows for evaluating a defense's skill in disguising their coverage and confusing the offense during pre-snap motion.

While our pre- and post-motion models are limited due to small sample size available, our HMM approach is modular and therefore our models can be replaced by more another model in the case of a richer data set. Combining both of these worlds could fully leverage the insights provided by the present data.
